Precision calculations of cosmic shear projections

### 2017-02-11 limber2.tex
Included Ludo's comments

\newcommand{\mk}[1]{{\bf\textcolor{blue}{#1}}}

@@

The tomographic shear power spectrum, in analogy to equation~(\ref{eq:C_ell_phi_Pm}), is defined by
%
\begin{equation}
  \left\langle _2\gamma^{}_{\ell m, i} \; {}_2\gamma^\ast_{\ell^\prime m^\prime, j} \right\rangle
    = \delta_{\ell \ell^\prime} \delta_{m m^\prime} C^\gamma_{ij}(\ell).
  \label{eq:C_ell_gamma}
\end{equation}
%
For a zero-curvature Universe, this is given by
%
\begin{align}
  C^\gamma_{ij}(\ell) = \frac 1 4 \ellbar^2(\ell, 2) \, C^\psi_{ij}(\ell)
                 = & \frac 2 \pi \, \ellbar^2(\ell, 2) \, \pref^2
                 \int_{0}^{\chi_{\rm lim}} \frac{\rm d \chi}{\chi} \frac{q_i(\chi)}{a(\chi)}
                \int_{0}^{\chi_{\rm lim}} \frac{\rm d \chi^\prime}{\chi^\prime}
                \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
                %\nonumber \\
                %& \times
                \int_0^\infty \frac{{\rm d} k}{k^2} \, P_{\rm m}(k, \chi, \chi^\prime) \,
                {\rm j}_\ell(k \chi) \, {\rm j}_\ell(k \chi^\prime) .
  \label{eq:C_ell_full}
\end{align}
%
The spherical pre-factor for the full projection power spectrum is $p(\ell) :=
\ellbar^2(\ell, 2)$, which will be modified under flat-sky and Limber
approximations below.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Flat-sky approximation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The majority of cosmic shear analyses have used lensing quantities approximated on a flat sky,
neglecting the sky curvature. This is a valid approach for past and current
survey areas with an extent less than 10 degrees. The correlation
functions from the observed shear have, however, been calculated using spherical
coordinates, and the effect on large scales has been shown to be significant \citep{FSHK08}.
Here we examine the effect of sky curvature on the theoretical models.

@@

To separate the $k$- and $\chi, \chi^\prime$-terms in equation~(\ref{eq:C_ell_full}), we
first approximate the matter power cross-spectrum between two
distances by the
geometric mean of the two involved distances \citep{2005PhRvD..72b3516C,2016arXiv161200770K},
%
\begin{equation}
 P_{\rm m}(k; \chi, \chi^\prime) = \sqrt{ P_{\rm m}(k; \chi) P_{\rm m}(k; \chi^\prime) } \, .
  \label{eq:P_geom_mean}
\end{equation}
%
With this, equation~(\ref{eq:C_ell_full}) is written as
%
\begin{align}
  C^\gamma_{ij}(\ell) \approx & \, \ellbar^2(\ell, 2) \, \pref^2
                \int_0^\infty \frac{{\rm d} k}{k^3} \,
                % P_{\rm m}(k) \,
                \int_{0}^\infty \frac{{\rm d} \chi}{\chi^{3/2}} \sqrt{P_{\rm m}(k; \chi)}
                \frac{q_i(\chi)}{a(\chi)} {\rm J}_{\ell+1/2}(k \chi)
                \nonumber \\
                 & \times
                \int_{0}^\infty \frac{{\rm d} \chi^\prime}{{\chi^\prime}^{3/2}}
                \sqrt{P_{\rm m}(k; \chi^\prime)} \frac{q_j(\chi^\prime)}{a(\chi^\prime)} {\rm J}_{\ell+1/2}(k \chi^\prime) \, .
  \label{eq:C_ell_full_Pk}
\end{align}
%
We replaced the upper limits of the distance integral with infinity. \mk{Comment on this.}
Note that this equation has a pre-factor $\ellbar^2(\ell, 2)$ corresponding to a spin-2 field, in contrast
to \cite{2008PhRvD..78l3506L} who show calculations for a scalar field.

Following \cite{2008PhRvD..78l3506L} we expand to third order
%
\begin{equation}
  \lim_{\varepsilon \rightarrow 0} \int_0^\infty {\rm d} x \, {\rm e}^{-\epsilon (x - \nu)} g(x) {\rm J}_\nu(x)
  %\approx g(\nu) - \frac 1 2 g^{\prime\prime}(\nu) - \frac \nu 6 g^{\prime\prime\prime}(\nu) 
  \approx g(\nu) - \frac 1 2 \left.\frac{{\rm d}^2 g}{{\rm d} x^2}\right|_{x=\nu}
                 - \frac \nu 6 \left.\frac{{\rm d}^3 g}{{\rm d} x^3}\right|_{x=\nu} ,
\end{equation}
%
with $g(x) = k^{-1} f(k, \chi)$, $\chi=x/k$, and its derivatives $g^{(n)}(x) =
k^{-1-n} f^{(n)}(k, \chi)$ for a given $k$. In our case the projection kernel is
%
\begin{equation}
  f(k, \chi) = \sqrt{P_{\rm m}(k; \chi)} \, a^{-1}(\chi) \chi^{-3/2} q(\chi).
  \label{eq:f_LA08}
\end{equation}
%
\mk{In the tomographic case, the index $i$ is added to $q$ and $f$.}
%
Replacing both distance integrals in equation~(\ref{eq:C_ell_full_Pk}) by their Taylor-expansions around $\nu = k \chi$ and $\nu = k \chi^\prime$,
respectively, yields
%
\begin{align}
  C^\gamma_{ij}(\ell) \approx & \, \ellbar^2(\ell, 2) \, \pref^2
    \int_0^\infty \frac{{\rm d} k}{k^3} \, k^{-2}
    \left[ f_i(k, \chi) - \frac{1}{2 k^2} f_i^{\prime\prime}(k, \chi)
      - \frac{\nu}{6 k^3} f_i^{\prime\prime\prime}(k, \chi) + \ldots \right]
    %\nonumber \\
    %& \times
    \left[ f_j(k, \chi) - \frac{1}{2 k^2} f_j^{\prime\prime}(k, \chi)
    - \frac{\nu}{6 k^3} f_j^{\prime\prime\prime}(k, \chi) + \ldots \right] \, .
  \label{eq:C_ell_limber2_dk}
\end{align}
%
Changing the integration to $\chi = \nu/k$ and collecting terms according to their $\nu$-dependence:
%
\begin{align}
  C^\gamma_{ij}(\ell) \approx & \, C^\gamma_{{\rm L1}, ij}(\ell) + C^\gamma_{{\rm L2}, ij}(\ell)
      \nonumber \\
    = & \frac{\ellbar^2(\ell, 2)}{\nu^4} \, \pref^2
    \int_0^\infty {\rm d} \chi \, \chi^3 \, % P_{\rm m}\left( k \right)
    %\Bigg\{
    \left\{
    (f_i f_j)(\nu/\chi, \chi)
    %\right.
      %\nonumber \\
    %&
    %\left.
     - \frac 1 {\nu^2} \left[ \frac{\chi^2}{2} \left( f^{}_i f^{\prime\prime}_j + f_i^{\prime\prime} f^{}_j \right)(\nu/\chi, \chi)
    + \frac{\chi^3}{6} \left( f^{}_i f^{\prime\prime\prime}_j + f^{\prime\prime\prime}_i f^{}_j \right)(\nu/\chi, \chi)
    \right]
    \right\} \, .
    %\Bigg\}
  \label{eq:C_ell_limber12_dr}
\end{align}
%
The first term corresponds to the well-known first-order Limber approximation
\citep{1953ApJ...117..134L,1992ApJ...388..272K}, which is widely used in
weak gravitational lensing. We retrieve the (spherical) standard expression by
inserting back the projection kernel equation~(\ref{eq:f_LA08}) and the time-dependent
power spectrum,
%
\begin{align}
  C^\gamma_{{\rm L1}, ij}(\ell) = & \, \frac{\ellbar^2(\ell, 2)}{\nu^4} \, \pref^2 \int {\rm d} \chi \frac{ q_i(\chi) q_j(\chi) }{a^2(\chi)}
  P_{\rm m}\left(\frac{\nu}{\chi}; \chi\right).
  \label{eq:C_ell_limber1}
\end{align}
%
In the Limber approximation, modes between structures at different epochs do
not contribute to the single line-of-sight integration.

@@

With this approximation, the \mk{$\chi$-independent term $P_{\rm m}(k)$}
factors out from equation~(\ref{eq:f_LA08}), and we define the separated kernel
function $f_{\rm s}$ as
%
\begin{equation}
  f_{\rm s}(\chi) = D_+(\chi) a^{-1}(\chi) \chi^{-3/2} q(\chi) \,
  \label{eq:f_LA08_s}
\end{equation}
%
The second-order Limber term of the shear power spectrum can then be simplified to
%
\begin{align}
  C^\gamma_{{\rm L2}, ij}(\ell) = & - \frac 1 {\nu^2} \, \frac{\ellbar^2(\ell, 2)}{\nu^4} \, \pref^2
    \int {\rm d} \chi \, \chi^{2} a^{-2} P_{\rm m}\left(\frac\nu\chi; \chi\right)
    %\nonumber \\
    %& \times
    \frac {a(\chi)}{2 D_+(\chi)} \chi^{3/2}
    \left[ q^{}_i f^{\prime\prime}_{{\rm s} j} + f^{\prime\prime}_{{\rm s}i} q^{}_j
      + \frac{\chi}{3} \left( q^{}_i f^{\prime\prime\prime}_{{\rm s}j} + f^{\prime\prime\prime}_{{\rm s}i} q^{}_j
      \right)
    \right](\chi)
  \label{eq:C_ell_limber2_dr_s}
\end{align}
%
We compute the numerical higher derivatives as follows. The functions $f_{\rm
s}(\chi)$ are fit as power laws with index $\approx 1.5$. The fit is
performed between $\chi_{\rm min} = 0.001 {\rm Mpc}/h$ and $\chi_{\rm max} =
500 \mbox{Mpc}/h$. At larger comoving distances the kernel decreases faster
than a power law, so we exclude this range from the fit. Even though the
derivaties are larger due to the steeper decline, the associated errors are
very small as these scales are down-weighed by the kernel function $f_{\rm
s}$ itself. At $\chi = 500 \, \mbox{Mpc}/h$ the filter function is less than
$10^{-4}$ of its value of unity at $1$\,Mpc$/h$.

@@

The extended flat-sky Limber approximation is readily derived from the
spherical case, by replacing the pre-factor $\ellbar^2(\ell, 2)$ with $\ell^4$,
%
\begin{align}
  P^\gamma_{{\rm L1}, ij}(\ell) = & \, \frac{\ell^4}{\nu^4} \, \pref^2 \int {\rm d} \chi \frac{ q_i(\chi) q_j(\chi) }{a^2(\chi)}
  P_{\rm m}\left(k = \frac{\nu}{\chi}; \chi\right);
  \label{eq:P_ell_limber1}
  \\
    P^\gamma_{{\rm L2}, ij}(\ell) = & - \frac 1 {\nu^2} \, \frac{\ell^4}{\nu^4} \, \pref^2
    \int {\rm d} \chi \, \chi^{2} a^{-2} P_{\rm m}\left(k = \frac\nu\chi; \chi\right)
    %\nonumber \\
    %& \times
    \frac{a(\chi)}{2 D^{-1}_+(\chi)} \chi^{3/2} \left[ q^{}_i f^{\prime\prime}_{{\rm s}j} + f^{\prime\prime}_{{\rm s}i} q^{}_j  
      + \frac{\chi}{3} \left( q^{}_i f^{\prime\prime\prime}_{{\rm s}j} + f^{\prime\prime\prime}_{{\rm s}i} q^{}_j
      \right)
    \right](\chi)
  \label{eq:P_ell_limber2}
\end{align}
%
Further approximations are commonly made for the pre-factor $p(\ell) = \ell^4/\nu^4$ and $\nu(\ell)$:
%
\begin{enumerate}
  \item $p(\ell) = 1$, this corresponds to $\nu(\ell) = \ell$, which is the \emph{standard} Limber approximation.
    Until recently, i.e.~for all pre-2014 CFHTLenS results, this was the approximation of choice.
  \item $p(\ell) = \ell^4/(\ell + 1/2)^4$. This corresponds to the \emph{extended} Limber
    approximation with $\nu(\ell) = \ell + 1/2$; however the following case is typically employed:
  \item $p(\ell) = 1$, with $\nu = \ell + 1/2$. This is
    a \emph{hybrid} between
    standard and extended Limber approximation, and was used in \cite{KiDS-450,joudaki/etal:2016}. As is shown
    later, this is a better approximation to the full projection than case (ii).
\end{enumerate}
%
\ch{Would this section be clearer if we had $p(\ell)$ in the equations instead of 
$\ell^4/\nu^4$?}

@@

\renewcommand{\baselinestretch}{1.5}
\begin{table}
\begin{centering}
  \caption{\label{tab:cases}The shear power spectrum approximations studied in this paper. 'ID' is the abbrevation used in the figures.
    The third column shows the power-spectrum pre-factor $p(\ell)$. }  
    \begin{tabular}{|l|l|c|c|c|p{0.21\textwidth}}
  \hline
  Case & ID & equation & $p(\ell)$ & $\nu(\ell)$ & Comment \\ \hline
  % 
  $1^{\rm st}$-order standard Limber, flat sky & L1Fl & (\ref{eq:P_ell_limber1}) + (i)
    & $1$ & $\ell$ & Used for pre-2014 CFHTLenS \\ \hline
  %
  $1^{\rm st}$-order extended Limber, flat sky & ExtL1Fl & (\ref{eq:P_ell_limber1}) + (ii)
    & $\frac{\ell^4}{(\ell+1/2)^4}$ & $\ell + \frac 1 2$ & Converges only with ${\cal O}(\ell^{-1})$ \\ \hline
  %
  $1^{\rm st}$-order extended Limber, hybrid, flat sky & ExtL1FlHyb & (\ref{eq:P_ell_limber1}) + (iii)
    & $1$ & $\ell + \frac 1 2$ & Used for post-2014 CFHTLenS, KiDS \\ \hline
  %
  $1^{\rm st}$-order extended Limber, spherical & ExtL1Sph & (\ref{eq:C_ell_limber1})
    & $\frac{\ellbar^2(\ell, 2)}{(\ell+1/2)^4}$ & $\ell+ \frac 1 2$ & \\ \hline
  %
  $2^{\rm st}$-order extended Limber, flat sky & ExtL2Fl &  (\ref{eq:P_ell_limber1}) + (\ref{eq:P_ell_limber2}) + (iii)
    & $\frac{\ell^4}{(\ell+1/2)^4}$ 
    & $\ell+\frac 1 2$ & Converges only with ${\cal O}(\ell^{-1})$ \\ \hline
  %
  $2^{\rm st}$-order extended Limber, hybrid, flat sky & ExtL2FlHyb &  (\ref{eq:P_ell_limber1}) + (\ref{eq:P_ell_limber2}) + (iii)
    & $1$ 
    & $\ell+ \frac 1 2$ & Best flat-sky approximation \\ \hline
  %
  $2^{\rm st}$-order extended Limber, spherical & ExtL2Sph & (\ref{eq:C_ell_limber1})+(\ref{eq:C_ell_limber2_dr_s})
    & $\frac{\ellbar^2(\ell, 2)}{(\ell+1/2)^4}$ & $\ell+\frac 1 2$ & Best approximation but mixing with flat-sky Hankel transform \\ \hline
  %
  Full spherical & FullSph & (\ref{eq:C_ell_full}) &
      $\ellbar^2(\ell, 2)$ & - & Correct projection \\ \hline
  %
  \end{tabular}

>>>>

\newcommand{\mk}[1]{{\bf\textcolor{blue}{#1}}}
\newcommand{\Label}[1]{\quad \mk{(\mbox{#1})}}
\newcommand{\LabelTxt}[1]{\mk{(\mbox{#1})}}

@@

The tomographic shear power spectrum, in analogy to equation~(\ref{eq:C_ell_phi_Pm}), is defined by
%
\begin{equation}
  \left\langle _2\gamma^{}_{\ell m, i} \; {}_2\gamma^\ast_{\ell^\prime m^\prime, j} \right\rangle
    = \delta_{\ell \ell^\prime} \delta_{m m^\prime} C^\gamma_{ij}(\ell).
  \label{eq:C_ell_gamma}
\end{equation}
%
For a zero-curvature Universe, this is given by
%
\begin{align}
  C^\gamma_{ij}(\ell) = \frac 1 4 \ellbar^2(\ell, 2) \, C^\psi_{ij}(\ell)
                 = & \frac 2 \pi \, \ellbar^2(\ell, 2) \, \pref^2
                 \int_{0}^{\chi_{\rm lim}} \frac{\rm d \chi}{\chi} \frac{q_i(\chi)}{a(\chi)}
                \int_{0}^{\chi_{\rm lim}} \frac{\rm d \chi^\prime}{\chi^\prime}
                \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
                %\nonumber \\
                %& \times
                \int_0^\infty \frac{{\rm d} k}{k^2} \, P_{\rm m}(k, \chi, \chi^\prime) \,
                {\rm j}_\ell(k \chi) \, {\rm j}_\ell(k \chi^\prime), \Label{FullSph}
  \label{eq:C_ell_full}
\end{align}
%
where we use the label `FullSph', see Table \ref{tab:cases} for a list of cases discussed in this work.
The spherical \mk{spin} pre-factor for the full projection \mk{shear} power
spectrum is $p(\ell) := \ellbar^2(\ell, 2)$, which will be modified under
flat-sky and Limber approximations below.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Flat-sky approximation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The majority of cosmic shear analyses have used \mk{the predicted} lensing
power spectrum approximated on a flat sky, neglecting the sky curvature. This
is a valid approach for past and current survey areas with an extent less than
10 degrees. \mk{To account for the sky curvature of the observed data}, the
shear correlation functions \mk{from observed galaxy ellipticities is now
routinely} computed using spherical coordinates, \mk{since projecting to a
Cartesian plane has been shown to cause significant biases of the two-point
funtion on large scales \citep{FSHK08}}. Here we examine the effect of sky
curvature on the theoretical \mk{models of the power spectrum, and the effect
on cosmological parameter inference (see Sect.~\ref{sec:cfhtlens}).}

@@

To separate the $k$- and $\chi, \chi^\prime$-terms in equation~(\ref{eq:C_ell_full}), we
first approximate the matter power cross-spectrum between two
distances by the
geometric mean of the two involved distances \citep{2005PhRvD..72b3516C,2016arXiv161200770K},
%
\begin{equation}
 P_{\rm m}(k; \chi, \chi^\prime) = \sqrt{ P_{\rm m}(k; \chi) P_{\rm m}(k; \chi^\prime) } \, .
  \label{eq:P_geom_mean}
\end{equation}
%
With this, equation~(\ref{eq:C_ell_full}) is written as
%
\begin{align}
  C^\gamma_{ij}(\ell) \approx & \, \ellbar^2(\ell, 2) \, \pref^2
                \int_0^\infty \frac{{\rm d} k}{k^3} \,
                % P_{\rm m}(k) \,
                \int_{0}^\infty \frac{{\rm d} \chi}{\chi^{3/2}} \sqrt{P_{\rm m}(k; \chi)}
                \frac{q_i(\chi)}{a(\chi)} {\rm J}_{\ell+1/2}(k \chi)
                %\nonumber \\
                 %& \times
                \int_{0}^\infty \frac{{\rm d} \chi^\prime}{{\chi^\prime}^{3/2}}
                \sqrt{P_{\rm m}(k; \chi^\prime)} \frac{q_j(\chi^\prime)}{a(\chi^\prime)} {\rm J}_{\ell+1/2}(k \chi^\prime) \, .
  \label{eq:C_ell_full_Pk}
\end{align}
%
We replaced the upper limits of the distance integral with infinity. \mk{Comment on this.}
Note that this equation has a pre-factor $\ellbar^2(\ell, 2)$ corresponding to a spin-2 field, in contrast
to \cite{2008PhRvD..78l3506L} who show calculations for a scalar field.

Following \cite{2008PhRvD..78l3506L} we expand to third order
%
\begin{equation}
  \lim_{\varepsilon \rightarrow 0} \int_0^\infty {\rm d} x \, {\rm e}^{-\epsilon (x - \nu)} g(x) {\rm J}_\nu(x)
  %\approx g(\nu) - \frac 1 2 g^{\prime\prime}(\nu) - \frac \nu 6 g^{\prime\prime\prime}(\nu) 
  \approx g(\nu) - \frac 1 2 \left.\frac{{\rm d}^2 g}{{\rm d} x^2}\right|_{x=\nu}
                 - \frac \nu 6 \left.\frac{{\rm d}^3 g}{{\rm d} x^3}\right|_{x=\nu} ,
\end{equation}
%
with $g(x) = k^{-1} f(k, \chi)$, $\chi=x/k$, and its derivatives $g^{(n)}(x) =
k^{-1-n} f^{(n)}(k, \chi)$ for a given $k$. In our case the projection kernel is
%
\begin{equation}
  f(k, \chi) = \sqrt{P_{\rm m}(k; \chi)} \, a^{-1}(\chi) \chi^{-3/2} q(\chi).
  \label{eq:f_LA08}
\end{equation}
%
\mk{In the tomographic case, the index $i$ is added to $q$ and $f$.}
%
Replacing both distance integrals in equation~(\ref{eq:C_ell_full_Pk}) by their Taylor-expansions around $\nu = k \chi$ and $\nu = k \chi^\prime$,
respectively, yields
%
\begin{align}
  C^\gamma_{ij}(\ell) \approx & \, \ellbar^2(\ell, 2) \, \pref^2
    \int_0^\infty \frac{{\rm d} k}{k^3} \, k^{-2}
    \left[ f_i(k, \chi) - \frac{1}{2 k^2} f_i^{\prime\prime}(k, \chi)
      - \frac{\nu}{6 k^3} f_i^{\prime\prime\prime}(k, \chi) + \ldots \right]
    %\nonumber \\
    %& \times
    \left[ f_j(k, \chi) - \frac{1}{2 k^2} f_j^{\prime\prime}(k, \chi)
    - \frac{\nu}{6 k^3} f_j^{\prime\prime\prime}(k, \chi) + \ldots \right] \, .
  \label{eq:C_ell_limber2_dk}
\end{align}
%
Changing the integration to $\chi = \nu/k$ and collecting terms according to their $\nu$-dependence:
%
\begin{align}
  C^\gamma_{ij}(\ell) \approx & \, C^\gamma_{{\rm L1}, ij}(\ell) + C^\gamma_{{\rm L2}, ij}(\ell)
      \nonumber \\
    = & \frac{\ellbar^2(\ell, 2)}{\nu^4} \, \pref^2
    \int_0^\infty {\rm d} \chi \, \chi^3 \, % P_{\rm m}\left( k \right)
    %\Bigg\{
    \left\{
    (f_i f_j)(\nu/\chi, \chi)
    %\right.
      %\nonumber \\
    %&
    %\left.
     - \frac 1 {\nu^2} \left[ \frac{\chi^2}{2} \left( f^{}_i f^{\prime\prime}_j + f_i^{\prime\prime} f^{}_j \right)(\nu/\chi, \chi)
    + \frac{\chi^3}{6} \left( f^{}_i f^{\prime\prime\prime}_j + f^{\prime\prime\prime}_i f^{}_j \right)(\nu/\chi, \chi)
    \right]
    \right\} \, .
    %\Bigg\}
  \label{eq:C_ell_limber12_dr}
\end{align}
%
The first term corresponds to the well-known first-order Limber approximation
\citep{1953ApJ...117..134L,1992ApJ...388..272K}, which is widely used in
weak gravitational lensing. We retrieve the (spherical) standard expression by
inserting back the projection kernel equation~(\ref{eq:f_LA08}) and the time-dependent
power spectrum,
%
\begin{align}
  C^\gamma_{{\rm L1}, ij}(\ell) = & \, \frac{\ellbar^2(\ell, 2)}{\nu^4} \, \pref^2 \int {\rm d} \chi \frac{ q_i(\chi) q_j(\chi) }{a^2(\chi)}
  P_{\rm m}\left(\frac{\nu}{\chi}; \chi\right) . \Label{ExtL1Sph}
  \label{eq:C_ell_limber1}
\end{align}
%
In the Limber approximation, modes between structures at different epochs do
not contribute to the single line-of-sight integration.

@@

With this approximation, the \mk{$\chi$-independent term $P_{\rm m}(k)$}
factors out from equation~(\ref{eq:f_LA08}), and we define the separated kernel
function $f_{\rm s}$ as
%
\begin{equation}
  f_{\rm s}(\chi) = D_+(\chi) a^{-1}(\chi) \chi^{-3/2} q(\chi) \,
  \label{eq:f_LA08_s}
\end{equation}
%
The second-order Limber term of the shear power spectrum can then be simplified to
%
\begin{align}
  C^\gamma_{{\rm L2}, ij}(\ell) = & - \frac 1 {\nu^2} \, \frac{\ellbar^2(\ell, 2)}{\nu^4} \, \pref^2
    \int {\rm d} \chi \, \chi^{2} a^{-2} P_{\rm m}\left(\frac\nu\chi; \chi\right)
    %\nonumber \\
    %& \times
    \frac {a(\chi)}{2 D_+(\chi)} \chi^{3/2}
    \left[ q^{}_i f^{\prime\prime}_{{\rm s} j} + f^{\prime\prime}_{{\rm s}i} q^{}_j
      + \frac{\chi}{3} \left( q^{}_i f^{\prime\prime\prime}_{{\rm s}j} + f^{\prime\prime\prime}_{{\rm s}i} q^{}_j
      \right)
    \right](\chi) \, . \Label{ExtL2Sph}
  \label{eq:C_ell_limber2_dr_s}
\end{align}
%
We compute the numerical higher derivatives as follows. The functions $f_{\rm
s}(\chi)$ are fit as power laws with index $\approx 1.5$. The fit is
performed between $\chi_{\rm min} = 0.001 {\rm Mpc}/h$ and $\chi_{\rm max} =
500 \mbox{Mpc}/h$. At larger comoving distances the kernel decreases faster
than a power law, so we exclude this range from the fit. Even though the
derivaties are larger due to the steeper decline, the associated errors are
very small as these scales are down-weighed by the kernel function $f_{\rm
s}$ itself. At $\chi = 500 \, \mbox{Mpc}/h$ the filter function is less than
$10^{-4}$ of its value of unity at $1$\,Mpc$/h$.

@@

The extended flat-sky Limber approximation is readily derived from the
spherical case (\ref{eq:C_ell_limber1}, \ref{eq:C_ell_limber2_dr_s}), by
replacing the pre-factor $\ellbar^2(\ell, 2)$ with $\ell^4$,
%
\begin{align}
  P^\gamma_{{\rm L1}, ij}(\ell) = & \, p(\ell) \, \pref^2 \int {\rm d} \chi \frac{ q_i(\chi) q_j(\chi) }{a^2(\chi)}
  P_{\rm m}\left(k = \frac{\nu}{\chi}; \chi\right);
  \label{eq:P_ell_limber1}
  \\
    P^\gamma_{{\rm L2}, ij}(\ell) = & - \frac 1 {\nu^2} \, p(\ell) \, \pref^2
    \int {\rm d} \chi \, \chi^{2} a^{-2} P_{\rm m}\left(k = \frac\nu\chi; \chi\right)
    %\nonumber \\
    %& \times
    \frac{a(\chi)}{2 D^{-1}_+(\chi)} \chi^{3/2} \left[ q^{}_i f^{\prime\prime}_{{\rm s}j} + f^{\prime\prime}_{{\rm s}i} q^{}_j  
      + \frac{\chi}{3} \left( q^{}_i f^{\prime\prime\prime}_{{\rm s}j} + f^{\prime\prime\prime}_{{\rm s}i} q^{}_j
      \right)
    \right](\chi)
  \label{eq:P_ell_limber2}
\end{align}
%
Further approximations are commonly made for the pre-factor $p(\ell) = \ell^4/\nu^4$ and $\nu(\ell)$:
%
\begin{enumerate}
  \item $p(\ell) = 1$, this corresponds to $\nu(\ell) = \ell$, which is the
    \emph{standard} Limber approximation \LabelTxt{L1Fl}
    Until recently, i.e.~for all pre-2014 CFHTLenS results, this was the approximation of choice.
    Note that we do not discuss the non-extended second-order Limber approximation here.
  %
  \item \Label{ExtL[1$|$2]Fl}: $p(\ell) = \ell^4/(\ell + 1/2)^4$. This corresponds to the \emph{extended} Limber
    approximation \LabelTxt{ExtL1Fl,ExtL2Fl} with $\nu(\ell) = \ell + 1/2$; however the following case is typically employed:
  %
  \item $p(\ell) = 1$, with $\nu = \ell + 1/2$. This is
    a \emph{hybrid} between
    standard and extended Limber approximation \Label{ExtL1FlHyb,ExtL2FlHyb}, and this case to first order
    was used in \cite{KiDS-450,joudaki/etal:2016}. As is shown
    later, this is a better approximation to the full projection than case (ii).
\end{enumerate}
%

@@

\renewcommand{\baselinestretch}{1.5}
\begin{table}
\begin{centering}
  \caption{\label{tab:cases}The shear power spectrum approximations studied in this paper. 'ID' is the \mk{label} used in the \mk{text and} figures.
    The third column shows the power-spectrum pre-factor $p(\ell)$. }  
    \begin{tabular}{|l|l|c|c|c|p{0.21\textwidth}}
  \hline
  Case & ID & equation & $p(\ell)$ & $\nu(\ell)$ & Comment \\ \hline
  % 
  $1^{\rm st}$-order standard Limber, flat sky & L1Fl & (\ref{eq:P_ell_limber1}) + (i)
    & $1$ & $\ell$ & Used for pre-2014 CFHTLenS \\ \hline
  %
  $1^{\rm st}$-order extended Limber, flat sky & ExtL1Fl & (\ref{eq:P_ell_limber1}) + (ii)
    & $\frac{\ell^4}{(\ell+1/2)^4}$ & $\ell + \frac 1 2$ & Converges only with ${\cal O}(\ell^{-1})$ \\ \hline
  %
  $1^{\rm st}$-order extended Limber, hybrid, flat sky & ExtL1FlHyb & (\ref{eq:P_ell_limber1}) + (iii)
    & $1$ & $\ell + \frac 1 2$ & Used for post-2014 CFHTLenS, KiDS \\ \hline
  %
  $1^{\rm st}$-order extended Limber, spherical & ExtL1Sph & (\ref{eq:C_ell_limber1})
    & $\frac{\ellbar^2(\ell, 2)}{(\ell+1/2)^4}$ & $\ell+ \frac 1 2$ & \\ \hline
  %
  $2^{\rm st}$-order extended Limber, flat sky & ExtL2Fl &  (\ref{eq:P_ell_limber1}) + (\ref{eq:P_ell_limber2}) + (iii)
    & $\frac{\ell^4}{(\ell+1/2)^4}$ 
    & $\ell+\frac 1 2$ & Converges only with ${\cal O}(\ell^{-1})$ \\ \hline
  %
  $2^{\rm st}$-order extended Limber, hybrid, flat sky & ExtL2FlHyb &  (\ref{eq:P_ell_limber1}) + (\ref{eq:P_ell_limber2}) + (iii)
    & $1$ 
    & $\ell+ \frac 1 2$ & Best flat-sky approximation \\ \hline
  %
  $2^{\rm st}$-order extended Limber, spherical & ExtL2Sph & (\ref{eq:C_ell_limber1})+(\ref{eq:C_ell_limber2_dr_s})
    & $\frac{\ellbar^2(\ell, 2)}{(\ell+1/2)^4}$ & $\ell+\frac 1 2$ & Best approximation but mixing with flat-sky Hankel transform \\ \hline
  %
  Full spherical & FullSph & (\ref{eq:C_ell_full}) &
      $\ellbar^2(\ell, 2)$ & - & Correct projection \\ \hline
  %
  \end{tabular}
